---
title: "ECONOMETRIA_R"
subtitle: 'Probability Theory'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.1) Random Variables and Probability distributions
### Probability distribution of discrete random variable
```{r}
#Muestro aleatorio discreto, size 1
sample(1:6,1)
#Graficando probabilidades 
prob<-rep(1/6,6)
plot(prob,main='Prob dis',xlab='outcome')
#Para la probabilidad acumulada
cum_prob<-cumsum((prob))
plot(cum_prob,main='Acu Prob dis',xlab='outcome')
```


### Bernoulli trials
```{r}
#Outcome pueden ser strings
sample(c("H", "T"), 1)
#Si denotamos a k como probabilidad de exito de optener H en bernuolli k~B(n,p)
dbinom(x=5,size=10,prob=0.5)
#Y si ahora queremos P(4<= K <=7 )
sum(dbinom(x=4:7,size=10,prob=0.5))
# Otra forma es con pbinom que toma prob. acum.
#ajustando vector de outcomes
k<-0:10
# assign the probabilities
probability <- dbinom(x = k,
                      size = 10, 
                      prob = 0.5)

# plot the outcomes against their probabilities
plot(x = k, 
     y = probability,
     main = "Probability Distribution Function") 
# compute cumulative probabilities
prob <- pbinom(q = k, 
               size = 10, 
               prob = 0.5)

# plot the cumulative probabilities
plot(x = k, 
     y = prob,
     main = "Cumulative Distribution Function") 
```


### Expected Value and the Mean
```{r}
# compute mean of natural numbers from 1 to 6
mean(1:6)

# set seed for reproducibility
set.seed(1)
# compute the sample mean of 10000 dice rolls
mean(sample(1:6, 
           10000, 
           replace = T))
#Varianza 
var(1:6)
```


### Probability distributions of continouos random variables
```{r}
# define functions
f <- function(x) 3 / x^4
g <- function(x) x * f(x)
h <- function(x) x^2 * f(x)
# compute area under the density curve
area <- integrate(f, 
                 lower = 1, 
                 upper = Inf)$value
area
# Calculando la esperanza
EX <- integrate(g,
                lower = 1,
                upper = Inf)$value
EX
# compute Var(X)
VarX <- integrate(h,
                  lower = 1,
                  upper = Inf)$value - EX^2 
VarX
## [1] 0.75
```


### The normal distribution
d for “density” - probability function / probability density function
p for “probability” - cumulative distribution function
q for “quantile” - quantile function (inverse cumulative distribution function)
r for “random” - random number generator
```{r}
# draw a plot of the N(0,1) PDF
curve(dnorm(x),
      xlim = c(-3.5, 3.5),
      ylab = "Density", 
      main = "Standard Normal Density Function") 
# compute denstiy at x=-1.96, x=0 and x=1.96
dnorm(x = c(-1.96, 0, 1.96))
# plot the standard normal CDF
curve(pnorm(x), 
      xlim = c(-3.5, 3.5), 
      ylab = "Density", 
      main = "Standard Normal Cumulative Distribution Function")
# define the standard normal PDF as an R function
f <- function(x) {
  1/(sqrt(2 * pi)) * exp(-0.5 * x^2)
}
# define a vector of reals
quants <- c(-1.96, 0, 1.96)
# compute densities
f(quants)
# integrate f()
integrate(f, 
          lower = -Inf, 
          upper = 1.337)
#El valor es la probabiliad de observar un z<1.337
#o 
pnorm(1.337)
# P(-1.96<z<1.96)=1-2xP(z(<-1.96)) Osea que Z se encuentre hasta dos ds de distancia
# compute the probability
1 - 2 * (pnorm(-1.96))
```


### The chi-squared distributions
```{r}
# plot the PDF
curve(dchisq(x, df = 3), 
      xlim = c(0, 10), 
      ylim = c(0, 1), 
      col = "blue",
      ylab = "",
      main = "p.d.f. and c.d.f of Chi-Squared Distribution, M = 3")

# add the CDF to the plot
curve(pchisq(x, df = 3), 
      xlim = c(0, 10),add = TRUE,
      col = "red")

# add a legend to the plot
legend("topleft", 
       c("PDF", "CDF"), 
       col = c("blue", "red"), 
       lty = c(1, 1))
# plot the density for M=1
curve(dchisq(x, df = 1), 
      xlim = c(0, 15), 
      xlab = "x", 
      ylab = "Density", 
      main = "Chi-Square Distributed Random Variables")

# add densities for M=2,...,7 to the plot using a 'for()' loop 
for (M in 2:7) {
  curve(dchisq(x, df = M),
        xlim = c(0, 15), 
        add = T, 
        col = M)
}

# add a legend
legend("topright", 
       as.character(1:7), 
       col = 1:7 , 
       lty = 1, 
       title = "D.F.")
# Aumento de varianza mientras aumento fd
```


### T Student Distribution

```{r}
curve(dnorm(x), 
      xlim = c(-4, 4), 
      xlab = "x", 
      lty = 2, 
      ylab = "Density", 
      main = "Densities of t Distributions")

# plot the t density for M=2
curve(dt(x, df = 2), 
      xlim = c(-4, 4), 
      col = 2, 
      add = T)

# plot the t density for M=4
curve(dt(x, df = 4), 
      xlim = c(-4, 4), 
      col = 3, 
      add = T)

# plot the t density for M=25
curve(dt(x, df = 25), 
      xlim = c(-4, 4), 
      col = 4, 
      add = T)

# add a legend
legend("topright", 
       c("N(0, 1)", "M=2", "M=4", "M=25"), 
       col = 1:4, 
       lty = c(2, 1, 1, 1))
#Vemos la convergencia de t a medida que aumentamos fd
```


### F distribution
```{r}
pf(2, 3, 13, lower.tail = F) # Prob que sea superior a 2, 2y13 son fd
# define coordinate vectors for vertices of the polygon
x <- c(2, seq(2, 10, 0.01), 10)
y <- c(0, df(seq(2, 10, 0.01), 3, 14), 0)

# draw density of F_{3, 14}
curve(df(x ,3 ,14), 
      ylim = c(0, 0.8), 
      xlim = c(0, 10), 
      ylab = "Density",
      main = "Density Function")

# draw the polygon
polygon(x, y, col = 30)



```




## 2.2)Random Sampling and the distribution of samples averages
```{r}
sum(sample(1:6, 2, replace = T))
# Vector of outcomes
S <- 2:12

# Vector of probabilities
PS <- c(1:6, 5:1) / 36

# Expectation of S
ES <- sum(S * PS) #notese * multiplica vectores

ES
# Variance of S
VarS <- sum(
  (S - c(ES))^2 * PS
  )
VarS
# divide the plotting area into one row with two columns
par(mfrow = c(1, 2))

# plot the distribution of S
barplot(PS, 
        ylim = c(0, 0.2), 
        xlab = "S", 
        ylab = "Probability", 
        col = "steelblue", 
        space = 0, 
        main = "Sum of Two Dice Rolls")

# plot the distribution of D 
probability <- rep(1/6, 6)
names(probability) <- 1:6

barplot(probability, 
        ylim = c(0, 0.2), 
        xlab = "D", 
        col = "steelblue", 
        space = 0, 
        main = "Outcome of a Single Dice Roll")
```
### Mean and Variance of the Sample Mean
```{r}
# set sample size and number of samples
n <- 10
reps <- 10000

# perform random sampling
samples <- replicate(reps, rnorm(n)) # 10 x 10000 sample matrix

# compute sample means
sample.avgs <- colMeans(samples)
# Plot the density histogram
hist(sample.avgs, 
     ylim = c(0, 1.4), 
     col = "steelblue" , 
     freq = F, 
     breaks = 20)

# overlay the theoretical distribution of sample averages on top of the histogram
curve(dnorm(x, sd = 1/sqrt(n)), 
      col = "red", 
      lwd = "2", 
      add = T)
# number of repititions
reps <- 10000

# set degrees of freedom of a chi-Square Distribution
DF <- 3 

# sample 10000 column vectors à 3 N(0,1) R.V.S
Z <- replicate(reps, rnorm(DF)) 

# column sums of squares
X <- colSums(Z^2)

# histogram of column sums of squares
hist(X, 
     freq = F, 
     col = "steelblue", 
     breaks = 40, 
     ylab = "Density", 
     main = "")

# add theoretical density
curve(dchisq(x, df = DF), 
      type = 'l', 
      lwd = 2, 
      col = "red", 
      add = T)
```


### Large Sample Aprroximations to Sampling Distributions

```{r}

# set seed
set.seed(1)

# set number of coin tosses and simulate
N <- 100
Y <- sample(0:1, N, replace = T)

# Calculate R_n for 1:N
S <- cumsum(Y)
R <- S/(1:N)

# Plot the path.
plot(R, 
     ylim = c(0.3, 0.7), 
     type = "l", 
     col = "steelblue", 
     lwd = 2, 
     xlab = "n", 
     ylab = "R_n",
     main = "Converging Share of Heads in Repeated Coin Tossing")

# Add a dashed line for R_n = 0.5
lines(c(0, N), 
      c(0.5, 0.5), 
      col = "darkred", 
      lty = 2, 
      lwd = 1)
#hacer juego de cambiar n, el 120-100> pareciese estar n
# subdivide the plot panel into a 2-by-2 array
par(mfrow = c(2, 2))

# set the number of repetitions and the sample sizes
reps <- 10000
sample.sizes <- c(5, 20, 75, 100)

# set seed for reproducibility
set.seed(123)

# vector of break numbers for histograms
b <- c(5, 10, 15, 30)

# outer loop (loop over the sample sizes)
  for (n in sample.sizes) {
    
    samplemean <- rep(0, reps) #initialize the vector of sample menas
    stdsamplemean <- rep(0, reps) #initialize the vector of standardized sample menas

# inner loop (loop over repetitions)   
    for (i in 1:reps) {
      x <- rbinom(n, 1, 0.5)
      samplemean[i] <- mean(x)
      stdsamplemean[i] <- sqrt(n)*(mean(x) - 0.5)/0.5
    }
    
# plot histogram and overlay the N(0,1) density in every iteration    
    hist(stdsamplemean, 
         col = "steelblue", 
         freq = FALSE, 
         breaks = 40,
         xlim = c(-3, 3), 
         ylim = c(0, 0.8), 
         xlab = paste("n =", n), 
         main = "")
    
    curve(dnorm(x), 
          lwd = 2, 
          col = "darkred", 
          add = TRUE)
  } 
```

